---
title: "JSON Vignette"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is JSON

Describe your type of data (JSON or XML). What is it, where does it get used, and why is it a good
way to store data? This should be detailed enough that someone that hasn’t seen that type of idea
would have a good idea what they are dealing with. You should link to references where applicable.

```{r cars}

```

## R packages

Discussion of possible packages/functions that are available for reading your type of data into R. Choose one and explain why you’ve chosen it.

```{r pressure, echo=FALSE}

```

## The data

### Description of dataset

Find a dataset of your type (JSON or XML) and describe where you found the data, how the data was
collected, what the variables are, etc.

### Reading in the data

Read in the data set describing the options your package’s functions allow.

### Exploratory Analysis

Perform basic exploratory data analysis that reveals a meaningful idea that you would reasonable want
to investigate further. Not all things reported need to show something meaningful (i.e. graphs that
show no relationship are fine) but you should end up with a solid lead that you would pursue further.

– At some point you should create a useful function(s) to do something meaningful with the data or
customize the way you read the data in.
– You should create a new variable.
– You should create some contingency tables and numeric summaries by some of your categorical
variables
– You should create some plots (at least a side-by-side bar plot, side-by-side box plots, and scatter
plots with coloring)


